Learning Actionable Analytics (with applications for reducing defects and reducing runtimes)

### 2015-05-05 how.tex
Updates from ShareLaTeX

 \documentclass[conference]{IEEEtran}
\usepackage{subfig}
\usepackage{wrapfig}
 \usepackage{amsmath}
 \usepackage{url}
 \usepackage{pifont}
 \usepackage{times}
\usepackage{rotating}
%\usepackage{balance} 
\usepackage{color, colortbl}
\usepackage{graphicx}
\usepackage{algorithmicx}

@@

\definecolor{Code}{rgb}{0,0,0}
\definecolor{Decorators}{rgb}{0.5,0.5,0.5}
\definecolor{Numbers}{rgb}{0.5,0,0}
\definecolor{MatchingBrackets}{rgb}{0.25,0.5,0.5}
\definecolor{Keywords}{rgb}{0,0,1}
\definecolor{self}{rgb}{0,0,0}
\definecolor{Strings}{rgb}{0,0.63,0}
\definecolor{Comments}{rgb}{0,0.63,1}
\definecolor{Backquotes}{rgb}{0,0,0}
\definecolor{Classname}{rgb}{0,0,0}
\definecolor{FunctionName}{rgb}{0,0,0}
\definecolor{Operators}{rgb}{0,0,0}
\definecolor{Background}{rgb}{1,1,1}

\lstnewenvironment{python}[1][]{
\lstset{
mathescape,
numbers=#1,
numberstyle=\footnotesize,
numbersep=0.5em,
xleftmargin=0em,
framextopmargin=2em,
framexbottommargin=2em,
showspaces=false,
showtabs=false,
showstringspaces=false,
tabsize=2,
% Basic
basicstyle=\ttfamily\small\setstretch{0.8},
backgroundcolor=\color{Background},
language=Python,
% Comments
commentstyle=\color{Comments}\slshape,
% Strings
stringstyle=\color{Strings},
morecomment=[s][\color{Strings}]{"""}{"""},
morecomment=[s][\color{Strings}]{'''}{'''},
% keywords
morekeywords={[1]import,from,class,def,for,while,if,is,in,elif,else,not,and,or,print,break,continue,return,True,False,None,access,as,,del,except,exec,finally,global,import,lambda,pass,print,raise,try,assert, dot, norm, zip, sorted},
keywordstyle={[1]\color{Code}\bfseries},
% additional keywords
morekeywords={[3]fastmap, project, furthest, split, WHERE,clusterer, getContours, envied, fWeight, nearestContour, projection, mutate, HERE, knn},
keywordstyle={[3]\color{Keywords}\bfseries},
morekeywords={[2]@invari},
keywordstyle={[2]\color{Decorators}\slshape},
emph={self},
emphstyle={\color{self}\slshape},
%
}}{}

@@

 \subsection{Motivation: Why ``HOW''?}
HOW was inspired by some recent work by Krall and Menzies and Davis on 
the GALE optimizer~\cite{krall14}. A well-known effect in machine learning (and, in computer vision
and compression~\cite{levina04}) is that the $n$ dimensions of a data set can be represented by a  much
lower-dimensional approximation. GALE finds those $m << n$ dimensions then mutates examples
towards the better end of each dimension $m_i$. Since it explores $m \ll n$ dimensions,
GALE terminates much faster than standard optimiser (e.g. for one large model, 4 minutes versus 7 hours~\cite{krall14}).
Better yet, GALE's rapdily generated  solutions are competitive (or better) than those found by 
slower optimizers.

A limitation with GALE is that it makes extensive use of
an underlying domain model to asses generations of newly mutated instances.  This is an issue since
in some domains, no such model exists or is readily accessible. For example, this whole paper
began when, after briefing a business client on GALE, they asked:
\begin{quote}
{\em ``But what if I do not have a model of my business?''}
\end{quote}
This is a legitimate concern.
Building and certifying a model  can  takes  long time, especially in software engineering. For  example, in previous work~\cite{me07f} we used models
developed by Boehm's group at the University of Southern California.
Some of those models took decades to develop and mature (from 1981~\cite{boehm81} to 2000~\cite{boehm00b}).  This is not an isolated example. One of us (Menzies) teaches a graduate class in automated model-based software engineering. 
Students in that class are asked to find (or build) a non-trivial model of
some software process (where ``non-trivial'' means something that processes
a  model at least an order of magnitude more complex than Brooks' Law). 
While the  number of such models has been growing in recent years\footnote{See the
catalogue at goo.gl/nv2AVK.}, 
the total number remains very small-- which is an observation consistent with
SE models typically require  much effort to build and certify.

@@

\subsection{Methods: How to Build ``HOW''?}
HOW is a variant of GALE that does not need a underlying domain model.

When designing HOW, our premises were:
\bi
\item
This is no domain model; (b)~we have some
labelled data and (b)~we seek minimal
disrupttion
HOW divide training into regions where the training
data has similar properties. Each region is then examined and divided into ``west'' and ``east''
where the region's ``west'' end has examples with worst performance stores than the east
end (mnemoc: west starts with ``w'' and is short for ``worst' and east starts with ``e;;]which is shirt
for excellent).so the ``e'' in

@@

PCA explores the $e$
eigenvectors of the correlation matrix between $d$
variables~\cite{pearson1901}.  Usually,
only a few $e \ll d$ eigenvectors are required
since multiple inter-related variables $i,j,k \subseteq d$ can be represented
by a single  eigenvector.
Also, noisy variables from $d$ are
ignored since they  do not contribute to the signal in the data.
Further,
redundant  variables  become (approximately) parallel lines
in $e$ space. For all such redundancies \mbox{$i,j \in d$}, we
can ignore $j$
since effects that change over $j$ also
change in the same way over $i$.

Algorithms like PDDP~\cite{boley98} use PCA to
recursively divide data into smaller regions. Each
step of that process requires an $O(N^2)$ analysis
of the data to compute the components~\cite{ilin10}. Worse, it requires an evaluation of every example.
Using WHERE, competitive results can be achieved
using an $O(2N)$ analysis.
WHERE uses a heuristic proposed by
Faloutsos and Lin~\cite{Faloutsos1995}
(which
Platt has
  shown computes a Nystr\"om
  approximation to the first component of
  PCA~\cite{platt05}).
In step one, WHERE inputs
$N$
examples $N_1,N_2,..$. Secondly, WHERE
picks any
point $N_i$ at random. Thirdly, WHERE
finds the point  {\em West}~$\in N$ that is
furthest\footnote{
 For this work, we use the standard Euclidean measure recommended for
instance-based reasoning by Aha et al.~\cite{aha91};
i.e. $\sqrt{\sum_i(x_i-y_i)^2}$ where $x_i,y_i$
are values normalized 0..1 for the range min..max.}
from $N_i$.
Finally, WHERE

$m \ll n$
dimension

>>>>

 \documentclass[conference]{IEEEtran}
\usepackage{subfig}
\usepackage{wrapfig}
 \usepackage{amsmath}
 \usepackage{url}
 \usepackage{pifont}
 %\usepackage{times}
\usepackage{rotating}
%\usepackage{balance} 
\usepackage{color, colortbl}
\usepackage{graphicx}
\usepackage{algorithmicx}

@@

\definecolor{Code}{rgb}{0,0,0}
\definecolor{Decorators}{rgb}{0.5,0.5,0.5}
\definecolor{Numbers}{rgb}{0.5,0,0}
\definecolor{MatchingBrackets}{rgb}{0.25,0.5,0.5}
\definecolor{Keywords}{rgb}{0,0,1}
\definecolor{self}{rgb}{0,0,0}
\definecolor{Strings}{rgb}{0,0.63,0}
\definecolor{Comments}{rgb}{0,0.63,1}
\definecolor{Comments}{rgb}{0.5,0.5,0.5}
\definecolor{Backquotes}{rgb}{0,0,0}
\definecolor{Classname}{rgb}{0,0,0}
\definecolor{FunctionName}{rgb}{0,0,0}
\definecolor{Operators}{rgb}{0,0,0}
\definecolor{Background}{rgb}{1,1,1}

\lstnewenvironment{python}[1][]{
\lstset{
mathescape,
numbers=#1,
numberstyle=\scriptsize,
numbersep=0.5em,
xleftmargin=0em,
framextopmargin=2em,
framexbottommargin=2em,
showspaces=false,
showtabs=false,
showstringspaces=false,
tabsize=2,
% Basic
basicstyle=\ttfamily\scriptsize\setstretch{0.8},
backgroundcolor=\color{Background},
language=Python,
% Comments
commentstyle=\color{Comments}\slshape,
% Strings
stringstyle=\color{Strings},
morecomment=[s][\color{Strings}]{"""}{"""},
morecomment=[s][\color{Strings}]{'''}{'''},
% keywords
morekeywords={[1]import,from,class,def,for,while,if,is,in,elif,else,not,and,or,print,break,continue,return,True,False,None,access,as,,del,except,exec,finally,global,import,lambda,pass,print,raise,try,assert, dot, norm, zip, sorted},
keywordstyle={[1]\color{Code}\bfseries},
% additional keywords
morekeywords={[3]fastmap, project, furthest, split, WHERE,clusterer, getContours, envied, fWeight, nearestContour, projection, mutate, HERE, knn},
keywordstyle={[3]\color{Keywords}\bfseries},
morekeywords={[2]@invari},
keywordstyle={[2]\color{Decorators}\slshape},
emph={self},
emphstyle={\color{self}\slshape},
%
}}{}

@@

 \subsection{Motivation: Why ``HOW''?}
HOW was inspired by certain  limits observed in 
the GALE optimizer~\cite{krall14}. 
A well-known effect in machine learning (and, in computer vision
and compression~\cite{levina04}) is that the $n$ dimensions of a data set can be represented by a  much
lower-dimensional approximation. GALE finds those $m << n$ dimensions then mutates examples
towards the better end of each dimension $m_i$. Since it explores $m \ll n$ dimensions,
GALE terminates much faster than standard optimiser (e.g. for one large model, 4 minutes versus 7 hours~\cite{krall14}).
Better yet, GALE's rapdily generated  solutions are competitive (or better) than those found by 
slower optimizers.

A limitation with GALE is that it makes extensive use of
an underlying domain model to asses generations of newly mutated instances.  This is an issue since
in some domains, no such model exists or is readily accessible. For example, this   paper
began when, after briefing a business client on GALE, they asked:
\begin{quote}
{\em ``But what if I do not have a model of my business?''}
\end{quote}
This is a legitimate concern.
Building and certifying a model  can  takes  long time, especially in software engineering. For  example, in previous work~\cite{me07f} we used models
developed by Boehm's group at the University of Southern California.
Some of those models took decades to develop and mature (from 1981~\cite{boehm81} to 2000~\cite{boehm00b}).  This is not an isolated example. One of us (Menzies) teaches a graduate class in automated model-based software engineering. 
Students in that class are asked to find (or build) a non-trivial model of
some software process (where ``non-trivial'' means something that processes
a  model at least an order of magnitude more complex than Brooks' Law). 
While the  number of such models has been growing in recent years\footnote{See the
catalogue at goo.gl/nv2AVK.}, 
the total number remains very small-- which is an observation consistent with
SE models typically require  much effort to build and certify.

@@

\subsection{Methods: How to Build ``HOW''?}

A general model for {\em predictors} is as follows. 
During testing, each test instance gains a  prediction 
by finding which region of the data it falls into.
During training, those regions are inferred from
divisions that ``chunk'' together regions with instances
with similar properties. These dividers can be categorical (as in the case of decision trees) or probabilistic (as in the
case of Naive Bayes) or learned via some distance minimization
process 
(as in the case of support vector machines) or generated
directly by some clustering algorithm.

Those same regions can be used for instance-based 
{\em planning}-- but instead of leaving test instances in the regions
found during training, we displace them towards regions
with better objective scores. 
Two important design decision with HOW 
is how to determine the {\em magnitude} and {\em direction} of the
displacement.

For HOW v1.0 (reported in this paper), we committed to a policy
of {\em minimal displacement}  since we do not wish to alarm developers and managers
with large scale change requests.
Note that minimal displacement makes HOW very different
to evolutionary optimizers.
Those algorithms, which include  
GALE, NSGA-II, NSGA-III, SPEA2, IBEA, PSO, DE and  MOEA/D~\cite{krall14,deb00a,zit02,zit04,%
deb14,Cui2005a,storn97,zhang07:TEC} can make large scale
displacements to the original population.

Accordingly, HOW v1.0  executes by:
\be
\item Dividing the $N$ training examples into many small clusters (each of size of , at most, $\sqrt{N}$);
\item
For a test instance that falls into cluster $C_i$,
never displace more than the boundary of that  cluster.
\ee
For the purposes of clustering, we use the WHERE recursive
descent procedure~\cite{me
WHERE is a spectral clusterer; i.e. it
executes over   synthesized dimensions discovered
by a 
principle components analysis (PCA)

Having  limiting the displacement {\em distance} to within one cluster,
there is still the issue of how to select the displacement 
{\em direction}.

Let the training and test instances comprise pair of independent and objective values. For example:
\bi 
\item
For the Seigmund data,
the objective score is the runtime of the test suite of some compiled program;
\item
For the Jureczko data, the objective score
is a boolean that is true if defects have been noted
for that class.
\ei
Let $X,Y$ be particular  training and test instance.
Further, 
let the training data be clustered and $X$
comes from a cluster $C_i$ that is near the test instance $Y$
and has objective scores better than those seen in $Y$.
We call $X$  the {\em exemplar} of $C_i$ since it
holds independent values that best characterize that cluster:
\bi
\item
For
instances with numeric objectives (e.g. the Seigmund data), 
the exemplar of a cluster is the instance
with the best objective score.
\item
For instances with discrete objectives (e.g. Jureczko data),
there is less differentiation between the objective scores.
Hence, for this kind of the data, the exemplar of a cluster is
an instance synthesized from the median values of the instances
in that cluster.
\ei
Given the above, we can now define a {\em plan} to be
a set of values to be added to a test instance in order
to reduce the distance between that instance and the exemplar: 
\[Y_j = Y_j + f*(X_j - Y_j)\]
In the above, $Y_j$ is an independent variable within the
test instance $Y$; $X$ is the exemplar found at training time;
and  $f$ is a constant that controls
how far we  move an instance (which is tuned for different data
sets).

A core design principle for HOW was ``minimal change'';
i.e. make very small changes to the current instance.
\ei
There are many ways to find those regions 
When designing HOW, our premises were:
\bi
\item
This is no domain model; (b)~we have some
labelled data and (b)~we seek minimal
disrupttion
HOW divide training into regions where the training
data has similar properties. Each region is then examined and divided into ``west'' and ``east''
where the region's ``west'' end has examples with worst performance stores than the east
end (mnemoc: west starts with ``w'' and is short for ``worst' and east starts with ``e;;]which is shirt
for excellent).so the ``e'' in

@@

PCA explores the $e$
eigenvectors of the correlation matrix between $d$
variables~\cite{pearson1901}.  
This is a useful approach since, usually,
only a few $e \ll d$ eigenvectors are required
since multiple inter-related variables $i,j,k \subseteq d$ can be represented
by a single  eigenvector.
Also, noisy variables from $d$ are
ignored since they  do not contribute to the signal in the data.
Further,
redundant  variables  become (approximately) parallel lines
in $e$ space. For all such redundancies \mbox{$i,j \in d$}, we
can ignore $j$
since effects that change over $j$ also
change in the same way over $i$.

Algorithms like PDDP~\cite{boley98} use PCA to
recursively cluster data into smaller regions. Each
step of that process requires an $O(N^2)$ analysis
of the data to compute the components~\cite{ilin10}. Worse, it requires an evaluation of every example.
Using WHERE, competitive results can be achieved
using an $O(2N)$ analysis.
WHERE uses a heuristic proposed by
Faloutsos and Lin~\cite{Faloutsos1995}
(which
Platt has
  shown computes a Nystr\"om
  approximation to the first component of
  PCA~\cite{platt05}).
Initially, WHERE inputs
$N$
instances $N_1,N_2,..$. Secondly, WHERE
picks any
instance $N_i$ at random. Thirdly, WHERE
finds the instance  {\em West}~$\in N$ that is
furthest from $N_i$. Finally,
if funds the point {\em East}~$\in N$ 
that is furthest from {\em West}. To computer these distances,
we use the standard Euclidean measure recommended for
instance-based reasoning by Aha et al.~\cite{aha91};
i.e. $\sqrt{\sum_i(x_i-y_i)^2}$ where $x_i,y_i$
are values normalized 0..1 for the range min..max.
